---
title: "Financial Analysis of long term care services for the elderly beings"
subtitle: "Proposal"
author: 
  - name: "Trailblazers"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(here)
library(dplyr)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

data <- read.csv(here("Long-Term_Care_Hospital-Provider_Data_Sep2024.csv"))

glimpse(data)



```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

## Source:

Data.CMS.gov

## How it was originally collected:

The dataset likely comes from CMS, with long-term care hospitals regularly reporting quality measures, which are audited and standardized, then published on Long-Term Care Hospital Compare for transparency and benchmarking.

## Brief description:

There are about 24826 rows and 16 columns in the dataset which accounts the data from 1960 to 2022 and the data sets show a detailed trend of various social expenditure on medicare from individual to government levels.

## Ethical Concerns:

Nothing Specific

## Questions:

1.  How do hospitals in different regions or states compare to the national benchmark in terms of performance scores?

2.  How is the distribution of long-term care hospitals across different states, counties, and CMS regions? Are there any regions with notably higher or lower numbers of these facilities?

3.  Has there been a change in hospital performance scores over the available time periods? Are there any seasonal or annual trends?

The two questions you want to answer.

If medical expenditure gets bigger as age increases, what are the sources of income and/or average income of that old age group?

How many percent of the population will be able to support themselves with proper medical treatment after retirement? what can individuals and governments do about it?

There are much options for categorical data

## Statement on why this question is important.

The aging population is a global trend for most the developed countries all over the world. asking and answering these research questions is valuable because they address critical issues related to aging, healthcare costs, and financial sustainability in retirement.

Predicting demographic trends over the next 15-20 years enables governments and businesses to proactively plan for shifts in labor markets, healthcare demand, and resource allocation. Finally, examining how many people can afford proper medical treatment after retirement highlights the need for both individuals and governments to take action, such as improving pension systems, enhancing healthcare access, and promoting financial preparedness, which are all essential for ensuring a healthier and more secure aging population.

## A description of the research topic along with a concise statement of your hypotheses on this topic.

1.  Medical expenditures for the elderly will significantly increase with age, and this group’s primary income sources will be insufficient to cover these expenses, leading to a financial strain fully.

2.  Demographic data will show a rapid increase in the elderly population in the US over the next 15-20 years, resulting in heightened demand for healthcare services, increased dependency ratios, and pressure on pension systems.

3.  A significant portion of the population will be unable to afford adequate medical treatment after retirement, necessitating policy interventions and increased individual financial planning to address this gap.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

1.  Trend Analysis of Total National Health Expenditures

This analysis could look at the total expenditures related to long-term care hospitals over time, comparing expenditure trends within and across states or CMS regions. Are there trends in costs that correlate with changes in performance scores, or are certain regions seeing more rapid expenditure increases? By examining the time-series data, this analysis could identify upward or downward trends in hospital expenditures, potentially shedding light on factors driving these shifts.

2.  Impact of Health Insurance on Expenditure

Although this dataset doesn't directly include insurance type or expenditures, exploring correlations between regions with high or low performance scores and general insurance coverage rates in those areas (from complementary datasets) could offer insights. For instance, regions with higher insurance coverage might show lower out-of-pocket expenditures, potentially affecting long-term care hospital utilization and quality. This can also lead to supplementary research on insurance coverage impact on expenditures across different regions.

3.  Out-of-Pocket Expenditures Analysis

If other sources on out-of-pocket costs are combined with this dataset, analysis could compare how out-of-pocket expenses vary for patients across hospitals with different performance scores or in different regions. By studying hospitals with different scores, one could examine if lower-scoring hospitals correlate with higher out-of-pocket expenses for patients, potentially impacting patient access to quality care in underperforming regions.

4.  Government vs. Private Expenditures

For regions or hospitals funded primarily through government programs (Medicare, Medicaid) versus those with more private funding, how do expenditures and performance compare? This would involve investigating whether government-supported regions correlate with higher or lower quality scores, thereby indicating the possible impact of funding sources on hospital performance. If supplemented by expenditure data, this could reveal the efficiency and outcomes tied to public vs. private healthcare funding.

5.  Expenditures on Different Services

With various Measure Codes representing different performance indicators (e.g., L_006_01_CI_LOWER, L_006_01_COMP_PERF), this dataset allows for exploring the expenditure or resource allocation patterns across services such as readmission rates, discharge times, or patient outcomes. Analyzing these measures can uncover whether high-expenditure areas show better outcomes across specific services and highlight which services are most correlated with increased quality

6.  Impact of Recessions on Health Expenditures

Recession periods (like the 2008 financial crisis or the economic downturn during COVID-19) could be aligned with expenditure and performance data to examine the resilience of long-term care hospitals. This analysis could explore whether recessions led to drops in performance or altered spending in particular regions. Regional analysis during these periods could highlight how economic stressors impact access to quality care and may affect performance scores.

7.  Forecasting Future Expenditures

Based on trends in performance and regional growth, forecasting models could project future costs associated with long-term care hospitals. By using past performance data and demographic trends, forecasts could estimate the future demand for long-term care services and anticipate which regions may face increasing expenditures. This could be useful for policymakers in planning for adequate resources and improving access to quality long-term care.
