---
title: "Financial analysis of medicare Services for the elderly population"
subtitle: "INFO 526 - Fall 2024 - Project 01"
author: 
  - name: "Trailblazers"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Financial Analysis of Medical Costs for the Elderly in the US"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(dplyr)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

url <- "https://raw.githubusercontent.com/INFO-526-F24/project-01-Trailblazers/refs/heads/main/data/NHE.csv"

data <- read.csv(url)

glimpse(data)
```

## Source:

Centers for Medicare & Medicaid Services

## How the data was originally collected:

The data are sourced from the **Services Annual Survey (SAS)** and the **Economic Census**, both conducted annually by the U.S. Census Bureau. Additionally, the data includes estimates from the **National Health Expenditure Accounts (NHEA)**, which are the official figures for total healthcare spending in the United States. Dating back to 1960, the NHEA provides annual measurements of U.S. expenditures on healthcare goods and services, public health activities, government administration, the net cost of health insurance, and investments related to healthcare. These comprehensive estimates offer a detailed view of the nation’s healthcare spending trends over time.

## 

## Brief description of the dataset:

The dataset contains 544 rows and 64 columns, spanning data from 1960 to 2022. It provides a detailed overview of various social expenditures on healthcare, covering contributions from individuals to government levels. This comprehensive dataset highlights trends in Medicare and other healthcare-related spending over time, offering insights into the financial dynamics of healthcare in the United States.

## Ethical Concerns:

Nothing Specific

## Questions:

1.  If medical expenditure increases with age, what are the primary sources of income for the elderly, and what is the average income for this age group?

2.  Given the U.S. population demographic trends, what predictions can be made for the next 15-20 years in terms of healthcare demand and financial sustainability for the aging population?

3.  What percentage of the population will be able to afford proper medical treatment after retirement? What actions can individuals and governments take to address this challenge?

## The two key questions we aim to address are

1.  As medical expenditure rises with age, what are the sources of income, and do the elderly have sufficient financial resources to cover healthcare costs?

2.  What percentage of the population will be able to afford adequate medical care post-retirement, and what steps can individuals and governments take to ensure access to necessary treatments?

## **Data Characteristics**

The dataset contains only numerical columns, with no categorical variables, and is well-structured with 543 rows across the dataset. All data is properly columnized, with each variable in its own column and each observation in its own row, ensuring the dataset is tidy and well-organized for analysis.

## Why these questions are important

The aging population is a significant global trend, particularly in developed countries. Exploring and addressing key research questions on this topic is crucial for understanding healthcare costs, financial sustainability in retirement, and, furthermore, social stability.

By analyzing demographic trends alongside financial projections of Medicare expenditures for the next 15-20 years, governments and businesses can better prepare for the rising healthcare costs associated with an aging population. Evaluating retirees’ ability to afford medical care emphasizes the urgent need for coordinated efforts from both the government and the private healthcare sector to improve access to care, affordability, and financial planning to meet the increasing demands of an aging population.

## **Research Topic and Hypotheses**

Medical expenditures for the elderly are expected to rise significantly with age, and the primary income sources for this group will likely be insufficient to fully cover the costs, leading to increasing financial strain.

Demographic data will predict a sharp rise in the elderly population in the U.S. over the next 15-20 years, which will result in greater demand for healthcare services, increased dependency ratios, and heightened pressure on pension and medicare systems.

A substantial portion of the population may be unable to afford adequate medical treatment after retirement, highlighting the need for policy interventions and enhanced individual financial planning to address this growing gap.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

1.  Trend Analysis of Total National Health Expenditures
2.  Impact of Health Insurance on Expenditure
3.  Out-of-Pocket Expenditures Analysis
4.  Government vs. Private Expenditures
5.  Expenditures on Different Services
6.  Impact of Recessions on Health Expenditures
7.  Forecasting Future Expenditures.
