---
title: "
Predicting Healthcare Service Demands Using Simulated Data for Advanced Analytics"
subtitle: "Proposal"
author: 
  - name: "Trailblazers"
    affiliations:
      - name: "College of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(here)
library(dplyr)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

data <- read.csv(here("data", "healthcare_dataset.csv"))

glimpse(data)



```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

## Source:

Kaggle

## How it was originally collected:

This dataset was synthetically generated using Python's Faker library. It was designed to mirror real-world healthcare records while ensuring that no sensitive or actual patient data is included, adhering strictly to privacy and ethical standards.

## Brief description:

-   **Number of Rows:** 55,500

-   **Number of Columns:** 15

-   **Columns Include:**

    -   Patient demographics (e.g., Name, Age, Gender)

    -   Clinical details (e.g., Blood Type, Medical Condition, Medication)

    -   Admission and discharge details (e.g., Date of Admission, Discharge Date, Admission Type)

    -   Billing and hospital information (e.g., Billing Amount, Hospital, Insurance Provider)

    -   Test Results (categories: Normal, Abnormal, Inconclusive)

#### Trends Present in the Dataset:

1.  Distribution of patient demographics, such as age and gender.

2.  Most common medical conditions and their associated admission types.

3.  Trends in test results based on patient demographics and medical conditions.

4.  Relationships between billing amounts, hospitals, and insurance providers.

## Ethical Concerns:

Nothing Specific

## Questions:

1.  What demographic factors influence test results (Normal, Abnormal, Inconclusive)? Importance: Identifying demographic trends can improve targeted healthcare services and personalized medicine.

2.  What is the relationship between medical conditions and admission types (Emergency, Elective, Urgent)? Importance: Understanding these patterns can help in hospital resource planning and patient care prioritization.

3.  What factors contribute most to higher billing amounts? Importance: Insights here can inform healthcare cost optimization and insurance claim processes.

The two questions you want to answer.

If medical expenditure gets bigger as age increases, what are the sources of income and/or average income of that old age group?

How many percent of the population will be able to support themselves with proper medical treatment after retirement? what can individuals and governments do about it?

There are much options for categorical data

## Statement on why this question is important.

Examining the impact of patient demographics and medical conditions on test results is vital for improving healthcare outcomes and advancing precision medicine. Insights from such research help identify patterns and disparities that influence diagnostic accuracy, enabling healthcare providers to deliver personalized care tailored to individual needs. This understanding also facilitates better resource allocation, ensuring critical care reaches high-risk populations effectively. Moreover, it enhances predictive modeling for early intervention, supports data-driven decision-making, and guides public health policies to address systemic inequities. By focusing on these relationships, the study contributes to the development of more efficient, equitable, and patient-centered healthcare systems.

## A description of the research topic along with a concise statement of your hypotheses on this topic.

Hypothesis: Demographics such as age, gender, and medical conditions significantly influence the outcomes of medical tests during healthcare admissions.

In addition to this baseline hypothesis, we can propose:

1.  Age Groups: Older age groups are more likely to have "Abnormal" test results due to comorbidities, whereas younger demographics might show a higher rate of "Normal" outcomes.

2.  Gender Influence: Certain medical conditions (e.g., cardiovascular diseases or diabetes) and test results may differ significantly between genders.

3.  Condition-Specific Outcomes: Patients with chronic conditions like hypertension or diabetes may exhibit a higher proportion of "Abnormal" results, while those admitted for minor ailments may lean toward "Normal" outcomes.

4.  Admission Type Correlation: Emergency admissions are hypothesized to show a higher likelihood of "Abnormal" results compared to elective or urgent admissions due to the critical nature of emergency cases.

### Expected Outcomes:

1.  Identification of high-risk demographic groups (e.g., older males with diabetes) for targeted interventions.

2.  Insights into resource allocation based on the likelihood of "Abnormal" or "Inconclusive" results in specific demographics.

3.  Improved predictive models for healthcare outcomes that incorporate demographic data, aiding in precision medicine and operational efficiency.

### Importance of This Study:

Personalized Medicine: Tailored interventions based on identified demographic patterns can enhance patient care.

Healthcare Efficiency: Understanding the links between demographics and test results can help hospitals optimize diagnostic resources.

Policy Making: The insights can inform public health policies, focusing on vulnerable populations for better health outcomes.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

1.  Data Preprocessing:

Handle categorical data by encoding variables like Gender, Medical Condition, Admission Type, and Test Results. Convert date columns (e.g., Date of Admission and Discharge Date) into appropriate formats and calculate the length of stay.

2.  Exploratory Data Analysis (EDA):

Visualize the distribution of Test Results across different demographics and medical conditions. Identify correlations between numerical features like Age, Billing Amount, and Room Number.

3.  Feature Engineering:

Create additional features, such as length of hospital stay and admission seasonality.

4.  Modeling:

Frame this as a multiclass classification problem to predict Test Results (Normal, Abnormal, Inconclusive). Use algorithms such as Logistic Regression, Random Forest, and XGBoost. Evaluate model performance using metrics like accuracy, precision, recall, and F1-score.

5.  Insights and Recommendations:

Summarize key factors influencing test results. Provide actionable insights for improving healthcare analytics and decision-making.
